No Replication Without Process Documentation:  An Assessment of Management

### 2020-04-10 R/MSE-manuscript.Rmd
Edit text through the discussion!

Management strategy evaluations typically do assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.  Data limitation is a frequent concern in fisheries management, and contributes to the degree of uncertainty present in fisheries management.  Next to accounting for uncertainty, accessing the influence of data quality and availablity, i.e.,  monitoring methodology was the most prevalent driver included in MSEs.  This likely points to a desire for the design and use of cost-effective fisheries monitoring plans, suggesting future MSEs may be able to learn from the set including monitoring methodology as a driver. Additionally, the use of value of information analyses could expand upon the benefits of MSE as a tool to direct scientific resources [@runge_2011; @moore_2012]. While discussions of fisheries research and management seem to be moving toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.

Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include peoples objectives.  These objectives are likely to include conservation, yield, economic, and social concerns, however the MSEs we reviewed don't cover all categories of objectives, as yield and economic objectives were not included in about one quarter of MSEs, and social objectives were absent from all but 2 of the 30 evaluations. While we are unable to know the cause for this with more explicit documentatin of MSE processes, our results may provide some clues.

Scientists led, and participated in most of the MSEs, while decision-makers, decision making institutions, and stakeholders were relatively infrequent leaders and participants in the process.  The lack of participation by stakeholders may result in less frequent inclusion of social objectives. Another factor may be that MSE models are likely constructed mainly by ecosystem modelers, who are likely to be more comfortable modeling the consequences of conservation and yield objectives than economic or social objectives.  Given that scientists seem to be providing the objectives in many MSE processes, they may be selecting objectives that they are more comfortable evaluating. There also appears to be some degree of borrowing occurring in which scientists use a set of generic fishery objectives when conducting a MSE.  Perhaps because their were few social objectives included in these MSE, there was little inclusion of experts to aid in the prediction of consequences, or consequence prediction techniques other than simulation modeling.

Participants that would be likely to expand the set of objectives considered in a MSE rarely participated in the MSEs we reviewed. That is, facilitators and decision analysts were rarely participants in MSE processes.  The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes [@Vente].  Facilitators and decision analysts also aid decision processes by helping to diagnose the problem, which can result in selecting more appropriate analysis techniques for the problem type.  Finally, the additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in some of the MSEs reviewed here [@williams2011; @smith2013; @kolody2008].

The relative dearth of explicit documentation we found in our review of MSE publications may hinder learning by MSE practitioners.  While the modeling was clearly documented, the MSE process steps were often inexplicit.  Without explicit documentation it is unclear why the analysis was structured as it was, why a particular set of drivers was selected, and why a set of objectives or alternatives were chosen.  Perhaps the most crucial stage to enable understanding a replication of a decision resulting from a MSE is a clear presentation of how trade-offs between the multiple objectives were evaluated, and what the optimal or chosen alternative was that resulted from the MSE process. Different decision makers can rationally look at the same prediction of consequences and select a different preferred alternative if they treat the importance of the objectives differently.  Therefore, MSE documentation often is not supporting replication or learning due to the frequent absence of trade-off and decision documentation.

The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis (Cummings, in prep). Utilizing the structured decision making approach eases the documentation process by breaking decision making into more discrete stages. There may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.  Where the expertise and budget is available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.

### Explore and expand upon this review
To aid future learning from the MSEs that have been conducted to date we produced a shiny application [@R-Shiny] that enables interactive results to accompany this article and entry of additional MSE information (<https://jonathancummings.shinyapps.io/MSEreview/>).  We hope that as MSE are conducted in the future they are entered here so that MSE practitioners can learn from each other.

>>>>

We found that management strategy evaluations do typically assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.  Data limitation is a frequent concern in fisheries management, and contributes to the degree of uncertainty present in fisheries management.  Next to accounting for uncertainty, accessing the influence of data quality and availablity, i.e.,  monitoring methodology was the most prevalent driver included in MSEs.  This likely points to a desire for the design and use of cost-effective fisheries monitoring plans, suggesting future MSEs may be able to learn from the set including monitoring methodology as a driver. Additionally, the use of value of information analyses could expand upon the benefits of MSE as a tool to direct scientific resources [@runge_2011; @moore_2012]. While discussions of fisheries research and management seem to be moving toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.

Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include objectives that reflect stakeholders' and society's values.  These objectives are likely to include conservation, yield, economic, and social concerns. As noted, MSEs are described as a tool to evaluate what management strategy best achieves objectives, however the MSEs we reviewed don't cover all categories of objectives. Yield and economic objectives were not included in about one quarter of MSEs, and broader social objectives were absent from all but 2 of the 30 evaluations. While we are unable to know the cause for this without more explicit documentation of MSE processes, our results may provide some clues.

Scientists led, and participated in most of the MSEs, while decision-makers, decision making institutions, and stakeholders were relatively infrequent leaders or participants in the process.  The lack of participation by stakeholders may result in less frequent inclusion of social objectives, as well as yield and economic objectives specific to a particular fishery. Another factor may be that MSE models are likely constructed mainly by ecosystem modelers, who are likely to be more comfortable modeling the consequences of conservation and yield objectives than economic or social objectives.  Given that scientists seem to be providing the objectives in many MSE processes, they may be selecting objectives that they are more comfortable evaluating. There also appears to be some degree of borrowing occurring in which scientists use a set of generic fishery objectives when conducting a MSE.  Perhaps because their were few social objectives included in these MSEs, there was little inclusion of experts to aid in the prediction of consequences, or consequence prediction techniques other than simulation modeling.

Participants that would be likely to expand the set of objectives considered in a MSE rarely participated in the MSEs we reviewed. That is, facilitators and decision analysts were rarely participants in MSE processes.  The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes [@Vente].  The additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in some of the MSEs reviewed here [@williams2011; @smith2013; @kolody2008]. Facilitators and decision analysts can also aid decision processes by helping to diagnose the problem, which can result in selecting more appropriate analysis techniques for the problem type, elicit a more complete and representative set of objectives, and select analysis tools to best address decision impediments to identify the optimal management strategy [@Cummings_inprep].

The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis, such as expanding the focus on problem framing at the outset of MSEs, eliciting fully representative sets of objectives, and utilizing multiple attribute utility theory to evaluate tradeoffs between objectives (Cummings, in prep). Utilizing the structured decision making approach also eases the documentation process by breaking decision making into more discrete stages. However, there may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.  Where the expertise and budget is available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.

The relative dearth of explicit documentation we found in our review of MSE publications may hinder learning by MSE practitioners.  While the modeling was clearly documented, the MSE process steps were often inexplicit.  Without explicit documentation it is unclear why the analysis was structured as it was, why a particular set of drivers was selected, and why a set of objectives or alternatives were chosen.  Perhaps the most crucial stage to enable understanding a replication of a decision resulting from a MSE is a clear presentation of how trade-offs between the multiple objectives were evaluated, and what the optimal or chosen alternative was that resulted from the MSE process. Different decision makers can rationally look at the same prediction of consequences and select a different preferred alternative if they treat the importance of the objectives differently.  Therefore, MSE documentation often is not fully supporting replication or learning due to the frequent absence of trade-off and decision documentation.

### Explore and expand upon this review
To aid learning from the documentation that has occurred, and provide ready access to future MSE documentation we produced a shiny application [@R-Shiny] to enable an interactive means to explore the MSE literature (<https://jonathancummings.shinyapps.io/MSEreview/>). The MSE review shiny application enables user to interactively explore the results presented in this article. The application includes additional analyses that may be of interest, such as an analysis of what journals MSEs have been published in, who is authoring MSEs, and more. Users may also filter by driver, location, species, or other characteristics of MSEs to explore a set of MSEs of interest and relevance to aid learning in the MSE practitioner community. We hope that as MSE are conducted they are entered here so that MSE practitioners can further learn from each other in the future.
